---
title: deriving ecological indices from lidar data second approach
author: KL
date: '2022-01-25'
slug: []
categories:
  - Assignment
tags:
  - task
subtitle: ''
description: ''
image: ''
toc: yes
output:
  blogdown::html_page:
    keep_md: yes
weight: 100
---

# Deriving ecological indices from LiDAR data (second approach)

## Research Question:

Is it possible to derive a suitable set of predictor variables from LiDAR 
data to obtain a reliable prediction of the microclimate parameters temperature and humidity?'

## Task:

- Read the ressources related to forest and identify those which you will use for attempting the task.
- Decide which algorithms and indices are adequate to answer the research question
- Apply and document this findings with on base of the scripts of the this unit experiences 

The Paper  "Compatibility of Aerial and Terrestrial LiDAR for
Quantifying Forest Structural Diversity" by LaRue et al. (2020) thematizes mainly the comparison of airial (ALS) and teresstrial (TLS) LiDAR techniques for quantifying forest structural diversity since Structural Diversity is, as they state, a "key feautrue of forest ecosystems" (LaRue, 2020, S.1).
Structural Diversity therefore also can be used to predict microclimate (ebd., S.1). Their findings "indicated that aerial LiDAR could be of use in quantifying broad-scale variation in structural diversity across macroscales (ebd., S.1). So according to LaRue et al. (2020), our Research question should be answered with yes, it is possible. In their supplementary material, LaRue et al. list many different metrics for structural diversity and how they are calculated in R. This will be usefull later on.

To find the right predictors to obtain a prediction of temperature and humidity, I additionally studied chapter 8 of "The lidR package by Jean-Romain Roussel, Tristan R.H. Goodbody, Piotr Tompalski 2021-01-15" and tried some of the mentioned metrics as follows. Also the review "Lidar data as indicators for forest biological diversity: a review" by Ida Marielle Mienna et al.(2018) helped to get an overview of different possible predictor variables.
 

## Setup:
```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```
```{r results='hide'} 
# activate envimaR package
require(envimaR)

# set root directory
rootDIR = "E:/edu/agis"

appendpackagesToLoad = c("lidR","forestr","lmom", "rLiDAR", "mapview", "raster", "rlas", "sp", "sf","tiff","raster", "rgl","future")

# define additional subfolders comment if not needed
appendProjectDirList =  c("data/lidar/","data/lidar/l_raster","data/lidar/l_raw","data/lidar/l_norm")

## define current projection (It is not magic you need to check the meta data or ask your instructor) 
## ETRS89 / UTM zone 32N
proj4 = "+proj=utm +zone=32 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
epsg_number = 25832

# load setup Skript
source(file.path(envimaR::alternativeEnvi(root_folder = rootDIR),"src/agis_setup.R"),echo = TRUE)
```

If I wanted to use a LAS file (not a catalog):
```{r}
#---- Get all *.las files of a folder into a list
las_files = list.files(envrmt$path_l_raw,
                       pattern = glob2rx("*.las"),
                       full.names = TRUE)

las = lidR::readLAS(las_files[1])
las

```

Creating LASCatalog and normalizing the data:
```{r eval=FALSE}
# clean data
future::plan(future::multisession)
ctg <- lidR::readLAScatalog(las_files[1])
lidR::projection(ctg) <- 25832
lidR::opt_chunk_size(ctg) = 100
lidR::opt_chunk_buffer(ctg) <- 5
lidR::opt_progress(ctg) <- FALSE
lidR::opt_laz_compression(ctg) <- TRUE
ctg@output_options$drivers$Raster$param$overwrite <- TRUE

#- height normalisation within the point cloud
  # source: https://github.com/Jean-Romain/lidR/wiki/Rasterizing-perfect-canopy-height-models
# calculate dem
  lidR::opt_output_files(ctg) <- paste0(envrmt$path_l_raster,"/","MOF_lidar_2018","/{ID}","_dsm")
  dem <- grid_terrain(ctg, res = 2,lidR::knnidw(k = 6L, p = 2))
  
#- normalize height (point cloud)
lidR::opt_output_files(ctg) <- paste0(envrmt$path_l_norm,"/{ID}","_norm_height")
ctg_n=lidR::normalize_height(ctg,lidR::tin())

```


Use only four tiles of the normalized catalog to minimize computing time:
```{r}
ctg_n <- readLAScatalog(file.path(envrmt$path_l_norm, "Ausschnitt_las"))

# reassign the projection
sp::proj4string(ctg_n) <- sp::CRS(proj4)

# visualize
ctg_n
mapview::mapview(ctg_n)
```

## Metrics:

The following metrics will be calculated: 
- DEM
- CHM
- standard metrics from lidR package
- mean height (of height normalized data)
- standard deviation and mean of intensity
- point density

Read in the same four tiles as for ctg_n, but of the digital elevation model (DEM) and create one raster by merging the four tiles:
```{r}
dem_small <- list.files( file.path(envrmt$path_l_raster,"MOF_lidar_2018", "Ausschnitt_tif"), pattern="*.tif")
dem_small

dem <- list()

for(i in 1:length(dem_small)) {
  dem[[i]] <- raster(paste0(envrmt$path_l_raster, "/MOF_lidar_2018/", "Ausschnitt_tif/", dem_small[i]))
}
dem

dem <- do.call(merge, dem)
plot(dem)

extent(ctg_n)
extent(dem)
```

CHM:
```{r}
# calculate the chm with the pitfree algorithm
chm = lidR::grid_canopy(ctg_n, 
                        2, 
                        pitfree(c(0,2,5,10,15), 
                        c(0,1), 
                        subcircle = 0.2))

# write it to tif
#raster::writeRaster(chm,file.path(envrmt$path_l_raster,"mof_chm_four_tiles.tif"),overwrite=TRUE) 

# visualize with mapview:
mapview::mapview(chm,
                   legend=TRUE, 
                   layer.name = "canopy height model",
                   col = mvTop(256))

# visualize with tmap:
tmap::tm_shape(chm) +  tmap::tm_raster(title = "CHM") +
  tm_grid() +
  tm_layout(legend.outside = TRUE)
```

Standard metrics:
```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# with the function stdmetrics, we can calculate many often used metrics (including f.e. height max and mean) at once: 
str_grid =grid_metrics(ctg_n, func = lidR::.stdmetrics, res = 2)
```

```{r}
plot(str_grid)
```

Mean and max height of height normalized data:
```{r echo=FALSE, results='hide'}
#cloud_metrics(ctg, func = ~mean(Z)) # does not work with ctg
#voxel_metrics(ctg, func = ~mean(Z)) # does not work with ctg

hmean <- grid_metrics(ctg_n, ~mean(Z), 2) # calculate mean at 10 m
hmax = grid_metrics(ctg_n, func = ~max(Z), 2)
hmin = grid_metrics(ctg_n, func = ~min(Z), 2)
```
```{r}
plot(hmean, col = height.colors(50))
plot(hmax, col = height.colors(50))
plot(hmin, col = height.colors(50))
```

Intensity:
```{r echo=FALSE, results='hide'}
# calculate the standard deviation and mean of intensity on grid level:
iSD=grid_metrics(ctg_n, func = sd(Intensity), 2)
imean= grid_metrics(ctg_n, func = mean(Intensity), 2)
```
```{r}
plot(iSD)
plot(imean)
```

Point density:
```{r echo=FALSE, results='hide'}
#might be an important ecological indice, since the structural density of the woods has large impact on f.e. plant growth which is directly connected to micro climate
density <- grid_metrics(ctg_n, ~length(Z)/4, 2) # calculate density
```
```{r}
plot(density, col = gray.colors(50,0,1)) # some plotting
```


Stack the calculated metrics:
```{r}

names(chm)="chm"
names(dem)="dem"
names(hmax) = "hmax"
names(hmean) = "hmean"
names(hmin)="hmin"
names(iSD)= "iSD"
names(imean)="imean"
names(density) = "density"

metricstack <- stack(chm,dem, hmax, hmean,hmin, iSD, imean, density)   
plot(metricstack)
```

## Literature:
LaRue, Elizabeth A. and Wagner, Franklin W. and Fei, Songlin and Atkins, Jeff W. and Fahey, Robert T. and Gough, Christopher M. and Hardiman, Brady S. (2020): Compatibility of Aerial and Terrestrial LiDAR for Quantifying Forest Structural Diversity. Remote Sensing Vol. 12.

Jean-Romain Roussel, 
Tristan R.H. Goodbody, Piotr Tompalski (2021): The lidR package.<https://jean-romain.github.io/lidRbook/>.

Ida Marielle Mienna, Katrine Eldegard, Ole Martin Bollandsås, Terje Gobakken1 and Hans Ole
Ørka (2018): Lidar data as indicators for forest biological diversity: a review.
